### 2017-04-24 caption/issues/07-11-09-ctt.tex

\iffalse
Date: Fri, 09 Nov 2007 22:19:11 +0100
From: =?UTF-8?B?IkNocmlzdGlhbiBFLiBCw7ZobWUi?= <hacks@cb8_replace_with_dot_org>
Organization: Wouldn't you like to know ?!!
User-Agent: Mozilla/5.0 (X11; U; Linux ppc; en-US; rv:1.8.0.13pre) Gecko/20070505 Iceape/1.0.9 (Debian-1.0.10~pre070720-0etch3+lenny1)
Newsgroups: comp.text.tex
Subject: figures, captions and skips
Message-ID: <4734cee5$0$13123$9b4e6d93@newsspool2.arcor-online.net>

Hi all,

while including figures/images of various widths, I noticed
that within the figure environment apparently the space/skip
generated between the caption and the included image changes
proportionally to the width of the image being included.  To
be precise, I would expect the code

\documentclass{book}

...

\usepackage{graphicx}
\usepackage{caption}
\captionsetup{position=bottom,font=small,labelfont=bf,format=hang}
\usepackage{subfig}

...

\begin{figure}
\centering
\includegraphics[keepaspectratio=true]{the-image.pdf}
\captionsetup{aboveskip=1em,belowskip=1em}
\caption[the title]{the caption}
\end{figure}

to place a caption below the image (which is assumed to be of
\textwidth width) using the same aboveskip amount that is being
generated for an image of half that width.  However, the code

...

\begin{figure}
\centering
\subfloat[][]{\includegraphics[keepaspectratio=true]{the-image.pdf}}
\captionsetup{aboveskip=1em,belowskip=1em}
\caption{...}
\end{figure}

...

does produce the intended aboveskip but with the drawback of a
subnumber on a one image figure which is somewhat undesirable.
Setting the global \abovecaptionskip had the same effect as
\captionsetup{aboveskip=1em}, ie, only producing an offset
bt not an absolute distance.

Is there a way to unconditionally/globally set the aboveskip
for a caption of a figure to a value that remains constant for
all images irrespective of their dimension ?

Thanks and regards,
Christian Boehme
\fi

\documentclass{book}

%...

\usepackage[demo]{graphicx}
\usepackage{caption}
\captionsetup{position=bottom,font=small,labelfont=bf,format=hang}
%\usepackage{subfig}

%...

\begin{document}

\begin{figure}
\centering
\includegraphics[keepaspectratio=true,width=0.5\linewidth]{the-image.pdf}
\captionsetup{aboveskip=1em,belowskip=1em}
\caption[the title]{the caption}
\end{figure}

\begin{figure}
\centering
\includegraphics[keepaspectratio=true,width=0.7\linewidth]{the-image.pdf}
\captionsetup{aboveskip=1em,belowskip=1em}
\caption[the title]{the caption}
\end{figure}

\begin{figure}
\centering
\includegraphics[keepaspectratio=true,width=1.0\linewidth]{the-image.pdf}
\captionsetup{aboveskip=1em,belowskip=1em}
\caption[the title]{the caption}
\end{figure}

\end{document}