### 2017-04-24 caption/issues/09-11-14.tex

\iffalse
Return-Path: <raphael.attie@mac.com>
Received: from compute1.internal (compute1.internal [10.202.2.41])
	 by store65m.internal (Cyrus v2.3.15-fmsvn20544-ebff1b04) with LMTPA;
	 Sat, 14 Nov 2009 11:01:05 -0500
X-Sieve: CMU Sieve 2.3
X-Spam-score: 0.0
X-Spam-hits: BAYES_00 -2.599, UNPARSEABLE_RELAY 0.001, BAYES_USED global
X-Spam-source: IP='81.169.145.136', Host='post.strato.de', Country='DE', FromHeader='com',
  MailFrom='com'
X-Spam-charsets: plain='US-ASCII', html='US-ASCII'
X-Resolved-to: sommerfee@f-m.fm
X-Mail-from: raphael.attie@mac.com
Received: from imap12.internal ([10.202.2.62])
  by compute1.internal (LMTPProxy); Sat, 14 Nov 2009 11:01:05 -0500
Received: from post.strato.de [81.169.145.136] by messagingengine.com
  with POP3 for sommerfee@f-m.fm; Sat, 14 Nov 2009 11:00:55 -0500
X-DeliveredInternalDate: Sat, 14 Nov 2009 15:19:48 +0100
X-LinkName: sommerfee.de
X-Envelope-From: <raphael.attie@mac.com>
X-Envelope-To: <ml@axel-sommerfeldt.de>
X-Delivery-Time: 1258208395
X-UID: 8863
X-RZG-FWD-BY: caption@sommerfee.de
Received: from RZmta-intern (client mail forwarder)
	by mailin.webmailer.de (blert mi9) (RZmta 22.1)
	for <ml@axel-sommerfeldt.de>; Sat, 14 Nov 2009 15:19:55 +0100 (MET)
X-RZG-CLASS-ID: mi
Received: from asmtpout012.mac.com ([17.148.16.87])
	by mailin.webmailer.de (blert mi9) (RZmta 22.1)
	with ESMTP id i040b4lAEEFtgF for <caption@sommerfee.de>;
	Sat, 14 Nov 2009 15:19:55 +0100 (MET)
MIME-version: 1.0
Content-type: multipart/alternative;
 boundary="Boundary_(ID_8vaiOHVH+3CE5q6BOg9/XQ)"
Received: from [192.168.2.101] ([82.82.127.216])
 by asmtp012.mac.com (Sun Java(tm) System Messaging Server 6.3-8.01 (built Dec
 16 2008; 32bit)) with ESMTPSA id <0KT300NZBRT01350@asmtp012.mac.com> for
 caption@sommerfee.de; Sat, 14 Nov 2009 06:19:53 -0800 (PST)
Message-id: <C6C43F24-B476-4F35-8A4C-011256DCBFEE@mac.com>
From: Raphael Attie <raphael.attie@mac.com>
To: caption@sommerfee.de
Subject: incompatibility of caption and university book-based document class,
 no-warning-issued
Date: Sat, 14 Nov 2009 15:19:48 +0100
X-Mailer: Apple Mail (2.936)

--Boundary_(ID_8vaiOHVH+3CE5q6BOg9/XQ)
Content-type: text/plain; charset=US-ASCII; format=flowed; delsp=yes
Content-transfer-encoding: 7BIT

Dear Axel Sommerfeldt,

I have read the documentation of the caption package, as I encountered  
a problem after the update from Mactex 2007 to Mactex 2009, thus using  
the version of the \caption package.

I am using a home-made "book"-based documentclass to write my Ph.D  
thesis, provided by the Max Planck Institute for Solar System Research  
in Germany. The class files are available there :
http://www.solar-system-school.de/thesis/thesis.zip

If needed, at the end of this email I've copied-pasted the content of  
thesis.cls that I'm using (together with the copy of my document), so  
that you don't have to download the zip file.

With the older version, Mactex 2007, this exact bit is working fine.  
Since the update, the caption package no longer works, only the  
\caption2 package does. However, it doesn't offer some of the  
improvements that the current \caption offers, and which I need, thus  
I would like to make the later work. For instance, with \caption2, I  
would not know how to put the whole caption in italic, and not just  
the "figure #:", whereas it is possible with the current version of  
the caption package.

The bug report I have is not the expected warning ( thus I thought I  
should report it) that you describe in your documentation :

Package caption Warning: Unsupported document class (or package)  
detected,
(caption) usage of the caption package is not recommended.
See the caption package documentation for explanation.

Instead, I have the following :

! Undefined control sequence.
\captionlabelfont@default ->\cph@font

As it worked before with 2007 version, and also work now with  
\caption2, I understand it comes from the thesis.cls not being  
compatible anymore.

Would you know a solution ?

Thank you for your help,

Raphael Attie
Ph.D Student
Max Planck Institute for Solar System Research
Max Planck Str. 2
37191 Katlenburg-Lindau
Germany

Here is a shortened version of my document.

------------------------
\fi

\documentclass[braunschweig]{09-11-14-thesis}

\usepackage{caption}
\usepackage[tight]{subfigure}
\usepackage{epstopdf}
\usepackage{fixltx2e}
\usepackage{float}
\usepackage{amsmath}
\usepackage{txfonts}

\begin{document}

\section{Tracking the photospheric flows}

blablabla

\begin{figure*}[h]
\centering
\includegraphics[width=1 \textwidth]{./figures/photosphere1.jpeg}
\caption{Image of the solar photosphere in continuum with MDI/SoHO,  
with a close-up taken at higher resolution by FG-SOT/Hinode. The  
latter shows the granulation.}  \label{granulation}
\end{figure*}

\end{document}
------------------------

-----------------------------

%% File: thesis.cls
%%
%% Author:
%% Dieter Schmitt
%% Max Planck Institute for Solar System Research
%% Max-Planck-Str. 2
%% D-37191 Katlenburg-Lindau
%% Germany
%%
%% Email: schmitt@mps.mpg.de
%%
%% For producing a PhD thesis for digital print and online publishing
%%
\NeedsTeXFormat{LaTeX2e}[1994/06/01]
\def\MyProvidesClass#1 [#2 #3 #4]
   {\ProvidesClass{#1}[#2 #3 #4]%
    \def\clsname{#1}\def\clsdate{#2}\def\clsversion{#3}\def\clsinfo{#4}}
\MyProvidesClass{thesis} [2008/07/22 1.8 (DS)]
%%
%% Version 1.2 (12 Jul 2005): change of Braunschweig title page, vec-- 
 >vect, ...
%% Version 1.3 (01 Oct 2006): change of Braunschweig title page,  
txfonts, pdfoutput
%% Version 1.4 (22 Dec 2006): options nottoc,notlof,notlot added to  
tocbibind
%% Version 1.5 (17 Apr 2007): change of Braunschweig title page
%% Version 1.6 (12 Oct 2007): Copernicus Publications and copyright
%% Version 1.7 (05 Mar 2008): change of copyright
%% Version 1.8 (22 Jul 2008): change of description
%%
\newif\if@goe
\newif\if@bs
\newif\if@goeorbs
\newif\if@sub
\newif\if@acc
\newif\if@print
\newif\if@file
\newif\if@fem
\newif\if@german
\newif\if@hyper
\DeclareOption{goettingen}{\@goetrue\@goeorbstrue\@subtrue}
\DeclareOption{braunschweig}{\@bstrue\@goeorbstrue\@subtrue}
\DeclareOption{submitted}{\@subtrue}
\DeclareOption{accepted}{\@acctrue\@subfalse}
\DeclareOption{print}{\@printtrue\@subfalse}
\DeclareOption{file}{\@filetrue\@subfalse}
\DeclareOption{female}{\@femtrue}
\DeclareOption{german}{\@germantrue}
\DeclareOption{hyper}{\@hypertrue}
\DeclareOption*{\PassOptionsToClass{\CurrentOption}{book}}
\ProcessOptions
\LoadClass[12pt]{book}
%%
\ifx\pdfoutput\undefined\else
    \setlength{\pdfpageheight}{29.7cm}
    \setlength{\pdfpagewidth}{21.0cm}
\fi
\setlength{\paperheight}{29.7cm}
\setlength{\paperwidth}{21.0cm}
\setlength{\textheight}{23.5cm}
\setlength{\textwidth}{15.0cm}
\setlength{\headheight}{2.0cm}
\setlength{\headsep}{0.8cm}
\setlength{\topskip}{12pt}
\setlength{\footskip}{30pt}
\setlength{\oddsidemargin}{0.46cm}
\setlength{\evensidemargin}{0.46cm}
\setlength{\topmargin}{-2.0cm}
%%
\ifx\pdfoutput\undefined
    \IfFileExists{graphicx.sty}{\RequirePackage[dvips]{graphicx}% LaTeX
    \DeclareGraphicsExtensions{.eps,.ps}}{}
    \IfFileExists{color.sty}{\RequirePackage{color}}{}
\else%
    \ifnum\pdfoutput=0
       \IfFileExists{graphicx.sty}{\RequirePackage[dvips]{graphicx}%  
LaTeX
       \DeclareGraphicsExtensions{.eps,.ps}}{}
       \IfFileExists{color.sty}{\RequirePackage{color}}{}
    \else
       \IfFileExists{graphicx.sty}{\RequirePackage[pdftex]{graphicx}%  
pdfLaTeX
       \DeclareGraphicsExtensions{.pdf,.png,.jpg}}{}
       \IfFileExists{color.sty}{\RequirePackage[pdftex]{color}}{}
       \pdfinfo{/Creator (thesis.cls version \clsversion)}
    \fi
\fi
%%
%\IfFileExists{txfonts.sty}{\RequirePackage{txfonts}}{}
\IfFileExists{natbib.sty}{\RequirePackage{natbib}}{}
\bibpunct{(}{)}{,}{a}{}{,}
%\setlength{\bibsep}{0pt}
\IfFileExists{fancyhdr.sty}{\RequirePackage{fancyhdr}}{}
\IfFileExists{tocbibind.sty}{\RequirePackage[nottoc,notlof,notlot] 
{tocbibind}}{}
\IfFileExists{fontenc.sty}{\RequirePackage[T1]{fontenc}}{}
\IfFileExists{url.sty}{\RequirePackage{url}}{}
\IfFileExists{amssymb.sty}{\RequirePackage{amssymb}}{}
%\IfFileExists{amsmath.sty}{\RequirePackage{amsmath}}{}
\IfFileExists{ifthen.sty}{\RequirePackage{ifthen}}{}
\IfFileExists{array.sty}{\RequirePackage{array}}{}
\IfFileExists{multicol.sty}{\RequirePackage{multicol}}{}
%%
\if@german
    \IfFileExists{german.sty}{\RequirePackage{german}}{}
    \def\captionsgerman{%
       \def\acknowname{Danksagung}%
       \def\refname{Literaturverzeichnis}%
       \def\abstractname{Zusammenfassung}%
%     \def\appendixname{Anhang}%
       \def\figurename{Abbildung}%
       \def\tablename{Tabelle}}
    \selectlanguage{german}
\else
    \def\captionsenglish{%
       \def\acknowname{Acknowledgement}%
       \def\refname{References}%
       \def\abstractname{Abstract}%
%     \def\appendixname{Appendix}%
       \def\figurename{Figure}%
       \def\tablename{Table}}
\fi
%%
\if@hyper\IfFileExists{hyperref.sty}{\RequirePackage{hyperref}}{}\fi
%%
\DeclareRobustCommand*{\vect}[1]{\ensuremath{%
\mathchoice{\mbox{\boldmath$\displaystyle#1$}}
            {\mbox{\boldmath$\textstyle#1$}}
            {\mbox{\boldmath$\scriptstyle#1$}}
            {\mbox{\boldmath$\scriptscriptstyle#1$}}}}
%%
\DeclareRobustCommand*{\chem}[1]{\ensuremath{\mathrm{#1}}}
\DeclareRobustCommand*{\unit}[1]{\def~{\,}\ensuremath{\mathrm{#1}}}
%%
\setcounter{topnumber}{2}
\def\topfraction{1.}
\setcounter{bottomnumber}{1}
\def\bottomfraction{1.}
\setcounter{totalnumber}{3}
\def\textfraction{0.}
\def\floatpagefraction{.7}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
%%
%\sloppy\clubpenalty4000\widowpenalty4000
%%
\def\town#1{\def\@town{#1}}               \def\@town{???}
\def\refereea#1{\def\@refa{#1}}           \def\@refa{???}
\def\refereeb#1{\def\@refb{#1}}           \def\@refb{???}
\def\submitteddate#1{\def\@subdate{#1}}   \def\@subdate{???}
\def\submittedyear#1{\def\@subyear{#1}}   \def\@subyear{\number\year}
\def\examinationdate#1{\def\@exdate{#1}}  \def\@exdate{???}
\def\publicationyear#1{\def\@pubyear{#1}} \def\@pubyear{\number\year}
\def\isbn#1{\def\@isbn{#1}}               \def\@isbn{???}
%%
\def\maketitle{%
\begin{titlepage}
\if@goe
    \vspace*{20mm}
    \begin{center}
    {\Huge\bfseries\@title\par}
    \vspace{30mm}
    {\large Dissertation\\[1mm]
    zur Erlangung des Doktorgrades\\[1mm]
    der Mathematisch-Naturwissenschaftlichen Fakult\"aten\\[1mm]
    der Georg-August-Universit\"at zu G\"ottingen
    \vfill
    vorgelegt von\\[1mm]
    {\Large\bfseries\@author}\\[1mm]
    aus \@town
    \par\vspace{30mm}
    G\"ottingen \space\@subyear}
    \end{center}
\fi
\if@bs
    \if@sub
       \vspace*{10mm}
       \begin{center}
       {\Huge\bfseries\@title\par}
       \vspace{20mm}
       {\large Der Fakult\"at f\"ur Elektrotechnik,  
Informationstechnik, Physik\\[1mm]
       der Technischen Universit\"at Carolo-Wilhelmina\\[1mm]
       zu Braunschweig\\[1mm]
       zur Erlangung des Grades \if@fem einer \else eines \fi \\[1mm]
       \if@fem Doktorin \else Doktors \fi der Naturwissenschaften\\[1mm]
       (Dr.rer.nat.)\\[1mm]
       eingereichte\\[1mm]
       Dissertation
       \par\vspace{30mm}
       von \@author\\[1mm]
       aus \@town\par}
       \end{center}
       \vfill
       {\large 1. Referentin oder Referent: \@refa\\[1mm]
       2. Referentin oder Referent: \@refb\\[1mm]
       eingereicht am: \@subdate\\[1mm]
       m\"undliche Pr\"ufung (Disputation) am:\par}
    \else
       \vspace*{20mm}
       \begin{center}
       {\Huge\bfseries\@title\par}
       \vspace{30mm}
       {\large Von der Fakult\"at f\"ur Elektrotechnik,  
Informationstechnik, Physik\\[1mm]
       der Technischen Universit\"at Carolo-Wilhelmina\\[1mm]
       zu Braunschweig\\[1mm]
       zur Erlangung des Grades \if@fem einer \else eines \fi \\[1mm]
       \if@fem Doktorin \else Doktors \fi der Naturwissenschaften\\[1mm]
       (Dr.rer.nat.)\\[1mm]
       genehmigte\\[1mm]
       Dissertation
       \par\vspace{30mm}
       von \@author\\[1mm]
       aus \@town\par}
       \end{center}
    \fi
\fi
\if@goeorbs\else\IfFileExists{firstpage.tex}{\include{firstpage}}
    {Please provide file firstpage.tex}\fi
\end{titlepage}
\newpage % second page
\thispagestyle{empty}
\if@goe
    \if@sub
       \vspace*{150mm}
       \noindent {\large D7\\[1mm]
       Referent: \@refa\\[1mm]
       Korreferent: \@refb\\[1mm]
       Tag der m\"undlichen Pr\"ufung: \par}
    \fi
    \if@print
       \noindent {\large {\bfseries Bibliografische Information Der  
Deutschen
       Bibliothek}\\[3mm]
       Die Deutsche Bibliothek verzeichnet diese Publikation in der  
Deutschen
       Nationalbibliografie; detaillierte bibliografische Daten sind im
       Internet \"uber {\tt http://dnb.ddb.de} abrufbar.
       \vfill
       \noindent D7\\[1mm]
       Referent: \@refa\\[1mm]
       Korreferent: \@refb\\[1mm]
       Tag der m\"undlichen Pr\"ufung: \@exdate\\[30mm]
       ISBN \@isbn\\[5mm]
       Copernicus Publications \@pubyear\\[1mm]
       http://publications.copernicus.org\\[1mm]
       \copyright\ \@author\\[5mm]
       Printed in Germany\par}
    \fi
    \if@file
       \IfFileExists{secondpage.tex}{\include{secondpage}}
       {Please provide file secondpage.tex}
    \fi
\fi
\if@bs
    \if@sub
       \null
    \fi
    \if@acc
       \vspace*{150mm}
       \noindent {\large 1. Referentin oder Referent: \@refa\\[1mm]
       2. Referentin oder Referent: \@refb\\[1mm]
       eingereicht am: \@subdate\\[1mm]
       m\"undliche Pr\"ufung (Disputation) am: \@exdate\par}
    \fi
    \if@print
       \noindent {\large {\bfseries Bibliografische Information Der  
Deutschen
       Bibliothek}\\[3mm]
       Die Deutsche Bibliothek verzeichnet diese Publikation in der  
Deutschen
       Nationalbibliografie; detaillierte bibliografische Daten sind im
       Internet \"uber {\tt http://dnb.ddb.de} abrufbar.
       \vfill
       \noindent 1. Referentin oder Referent: \@refa\\[1mm]
       2. Referentin oder Referent: \@refb\\[1mm]
       eingereicht am: \@subdate\\[1mm]
       m\"undliche Pr\"ufung (Disputation) am: \@exdate\\[30mm]
       ISBN \@isbn\\[5mm]
       Copernicus Publications \@pubyear\\[1mm]
       http://publications.copernicus.org\\[1mm]
       \copyright\ \@author\\[5mm]
       Printed in Germany\par}
    \fi
    \if@file
       \IfFileExists{secondpage.tex}{\include{secondpage}}
       {Please provide file secondpage.tex}
    \fi
\fi
\if@goeorbs\else\IfFileExists{secondpage.tex}{\include{secondpage}}
{Please provide file secondpage.tex}\fi
} % end of maketitle definition
%%
\newenvironment{cfig}{\begin{figure}[t]\begin{center}}{\end{center} 
\end{figure}}
%%
\def\chaptername{}
\def\appendixname{}
%%
\pagestyle{fancy}
\renewcommand{\chaptermark}[1]{\markboth{\thechapter\hskip2ex #1}
    {\thechapter\hskip2ex #1}}
\renewcommand{\sectionmark}[1]{\markright{\thesection\hskip2ex #1}}
\fancyhf{}
\fancyhead[LE]{\nouppercase{\leftmark}}
\fancyhead[RO]{\nouppercase{\rightmark}}
\fancyfoot[LE,RO]{\thepage}
\renewcommand{\footrulewidth}{0pt}
\fancypagestyle{plain}{\fancyhf{}\fancyfoot[LE,RO]{\thepage}%
    \renewcommand{\headrulewidth}{0pt}\renewcommand{\footrulewidth} 
{0pt}}
%%
% avoids headings on otherwise empty pages before a new chapter
\def\cleardoublepage{\clearpage\if@twoside\ifodd\c@page\else
    \hbox{}\thispagestyle{empty}\newpage
    \if@twocolumn\hbox{}\newpage\fi\fi\fi}
%%
%\def\@schapter#1{\addcontentsline{toc}{chapter}{#1} % for toc
%   \@mkboth{#1}{#1}                                 % headings
%   \if@twocolumn\@topnewpage[\@makeschapterhead{#1}]
%   \else\@makeschapterhead{#1}\@afterheading\fi}
%
% problem: adds also toc to toc
%%
\def\@makechapterhead#1{\vspace*{50\p@}{\parindent\z@\raggedright
    \normalfont\interlinepenalty\@M\Huge\bfseries{%
    \@hangfrom{\hskip\z@\relax\thechapter\hskip2ex}#1\@@par}
    \par\nobreak\vskip40\p@}}
%%
%% End of file `thesis.cls'.

--Boundary_(ID_8vaiOHVH+3CE5q6BOg9/XQ)
Content-type: text/html; charset=US-ASCII
Content-transfer-encoding: quoted-printable

<html><body style=3D"word-wrap: break-word; -webkit-nbsp-mode: space; =
-webkit-line-break: after-white-space; "><div>Dear Axel =
Sommerfeldt,</div><div><br></div><div>I have read the documentation of =
the caption package, as I encountered a problem after the update from =
Mactex 2007 to Mactex 2009, thus using the version of the \caption =
package.</div><div><br></div><div>I am using a home-made "book"-based =
documentclass to write my Ph.D thesis, provided by the Max Planck =
Institute for Solar System Research in Germany. The class files are =
available there :&nbsp;</div><div><a =
href=3D"http://www.solar-system-school.de/thesis/thesis.zip">http://www.so=
lar-system-school.de/thesis/thesis.zip</a></div><div><br></div><div>If =
needed, at the end of this email I've copied-pasted the content of =
thesis.cls that I'm using (together with the copy of my document), so =
that you don't have to download the zip =
file.</div><div><br></div><div>With the older version, Mactex 2007, this =
exact bit is working fine. Since the update, the caption package no =
longer works, only the \caption2 package does. However, it doesn't offer =
some of the improvements that the current \caption offers, and which I =
need, thus I would like to make the later work. For instance, with =
\caption2, I would not know how to put the whole caption in italic, and =
not just the "figure #:", whereas it is possible with the current =
version of the caption package.</div><div><br></div><div>The bug report =
I have is not the expected warning (&nbsp;thus I thought I should report =
it) that you describe in your documentation =
:&nbsp;</div><div><br></div><div><div><i>Package caption Warning: =
Unsupported document class (or package) detected,</i></div> =
<div><i>(caption) usage of the caption package is not =
recommended.</i></div> <div><i>See the caption package documentation for =
explanation.</i></div></div><br><div>Instead, I have the following =
:&nbsp;</div><div><br></div><div>! Undefined control =
sequence.<br>\captionlabelfont@default =
-&gt;\cph@font</div><div><br></div><div>As it worked before with 2007 =
version, and also work now with \caption2, I understand it comes from =
the thesis.cls not being compatible =
anymore.&nbsp;</div><div><br></div><div>Would you know a solution =
?</div><div><br></div><div>Thank you for your =
help,</div><div><br></div><div><br></div><div><div>Raphael =
Attie</div><div><div><div>Ph.D Student</div><div>Max Planck Institute =
for Solar System Research</div><div>Max Planck Str. 2</div><div>37191 =
Katlenburg-Lindau</div><div>Germany</div><div><br></div></div></div></div>=
<div><br></div><div><div><br></div><div><br></div><div><br></div><div>Here=
 is a shortened version of my =
document.&nbsp;</div><div><br></div></div><div>------------------------</d=
iv><div><br></div><div><div>\documentclass[braunschweig]{thesis}</div><div=
><br></div><div>\usepackage{caption}</div><div>\usepackage[tight]{subfigur=
e}</div><div>\usepackage{epstopdf}</div><div>\usepackage{fixltx2e}</div><d=
iv>\usepackage{float}</div><div>\usepackage{amsmath}</div><div>\usepackage=
{txfonts}</div><div><br></div><div>\begin{document}</div><div><br></div><d=
iv>\section{Tracking the photospheric =
flows}</div><div><br></div><div>blablabla&nbsp;</div><div><br></div><div>\=
begin{figure*}[h]</div><div>\centering</div><div>\includegraphics[width=3D=
1 \textwidth]{./figures/photosphere1.jpeg}</div><div>\caption{Image of =
the solar photosphere in continuum with MDI/SoHO, with a close-up taken =
at higher resolution by FG-SOT/Hinode. The latter shows the =
granulation.} &nbsp;\label{granulation}</div><div>\end{figure*}<span =
class=3D"Apple-tab-span" style=3D"white-space:pre">	=
</span></div><div><br></div><div>\end{document}</div>---------------------=
---</div><div><br></div><div><br></div><div><br></div>--------------------=
---------<div><br></div><div><div>%% File: =
thesis.cls</div><div>%%</div><div>%% Author:</div><div>%% Dieter =
Schmitt</div><div>%% Max Planck Institute for Solar System =
Research</div><div>%% Max-Planck-Str. 2</div><div>%% D-37191 =
Katlenburg-Lindau</div><div>%% Germany</div><div>%%</div><div>%% Email: =
<a =
href=3D"mailto:schmitt@mps.mpg.de">schmitt@mps.mpg.de</a></div><div>%%</di=
v><div>%% For producing a PhD thesis for digital print and online =
publishing</div><div>%%&nbsp;</div><div>\NeedsTeXFormat{LaTeX2e}[1994/06/0=
1]</div><div>\def\MyProvidesClass#1 [#2 #3 =
#4]</div><div>&nbsp;&nbsp;{\ProvidesClass{#1}[#2 #3 =
#4]%</div><div>&nbsp;&nbsp; =
\def\clsname{#1}\def\clsdate{#2}\def\clsversion{#3}\def\clsinfo{#4}}</div>=
<div>\MyProvidesClass{thesis} [2008/07/22 1.8 =
(DS)]</div><div>%%</div><div>%% Version 1.2 (12 Jul 2005): change of =
Braunschweig title page, vec--&gt;vect, ...</div><div>%% Version 1.3 (01 =
Oct 2006): change of Braunschweig title page, txfonts, =
pdfoutput</div><div>%% Version 1.4 (22 Dec 2006): options =
nottoc,notlof,notlot added to tocbibind</div><div>%% Version 1.5 (17 Apr =
2007): change of Braunschweig title page</div><div>%% Version 1.6 (12 =
Oct 2007): Copernicus Publications and copyright</div><div>%% Version =
1.7 (05 Mar 2008): change of copyright</div><div>%% Version 1.8 (22 Jul =
2008): change of =
description</div><div>%%</div><div>\newif\if@goe</div><div>\newif\if@bs</d=
iv><div>\newif\if@goeorbs</div><div>\newif\if@sub</div><div>\newif\if@acc<=
/div><div>\newif\if@print</div><div>\newif\if@file</div><div>\newif\if@fem=
</div><div>\newif\if@german</div><div>\newif\if@hyper</div><div>\DeclareOp=
tion{goettingen}{\@goetrue\@goeorbstrue\@subtrue}</div><div>\DeclareOption=
{braunschweig}{\@bstrue\@goeorbstrue\@subtrue}</div><div>\DeclareOption{su=
bmitted}{\@subtrue}</div><div>\DeclareOption{accepted}{\@acctrue\@subfalse=
}</div><div>\DeclareOption{print}{\@printtrue\@subfalse}</div><div>\Declar=
eOption{file}{\@filetrue\@subfalse}</div><div>\DeclareOption{female}{\@fem=
true}</div><div>\DeclareOption{german}{\@germantrue}</div><div>\DeclareOpt=
ion{hyper}{\@hypertrue}</div><div>\DeclareOption*{\PassOptionsToClass{\Cur=
rentOption}{book}}</div><div>\ProcessOptions</div><div>\LoadClass[12pt]{bo=
ok}</div><div>%%</div><div>\ifx\pdfoutput\undefined\else</div><div>&nbsp;&=
nbsp; \setlength{\pdfpageheight}{29.7cm}</div><div>&nbsp;&nbsp; =
\setlength{\pdfpagewidth}{21.0cm}</div><div>\fi</div><div>\setlength{\pape=
rheight}{29.7cm}</div><div>\setlength{\paperwidth}{21.0cm}</div><div>\setl=
ength{\textheight}{23.5cm}</div><div>\setlength{\textwidth}{15.0cm}</div><=
div>\setlength{\headheight}{2.0cm}</div><div>\setlength{\headsep}{0.8cm}</=
div><div>\setlength{\topskip}{12pt}</div><div>\setlength{\footskip}{30pt}<=
/div><div>\setlength{\oddsidemargin}{0.46cm}</div><div>\setlength{\evensid=
emargin}{0.46cm}</div><div>\setlength{\topmargin}{-2.0cm}</div><div>%%</di=
v><div>\ifx\pdfoutput\undefined</div><div>&nbsp;&nbsp; =
\IfFileExists{graphicx.sty}{\RequirePackage[dvips]{graphicx}% =
LaTeX</div><div>&nbsp;&nbsp; =
\DeclareGraphicsExtensions{.eps,.ps}}{}</div><div>&nbsp;&nbsp; =
\IfFileExists{color.sty}{\RequirePackage{color}}{}</div><div>\else%</div><=
div>&nbsp;&nbsp; \ifnum\pdfoutput=3D0</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\IfFileExists{graphicx.sty}{\RequirePackage[dvips]{graphicx}% =
LaTeX</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\DeclareGraphicsExtensions{.eps,.ps}}{}</div><div>&nbsp;&nbsp; =
&nbsp; =
&nbsp;\IfFileExists{color.sty}{\RequirePackage{color}}{}</div><div>&nbsp;&=
nbsp; \else</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\IfFileExists{graphicx.sty}{\RequirePackage[pdftex]{graphicx}% =
pdfLaTeX</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\DeclareGraphicsExtensions{.pdf,.png,.jpg}}{}</div><div>&nbsp;&nbsp;=
 &nbsp; =
&nbsp;\IfFileExists{color.sty}{\RequirePackage[pdftex]{color}}{}</div><div=
>&nbsp;&nbsp; &nbsp; &nbsp;\pdfinfo{/Creator (thesis.cls version =
\clsversion)}</div><div>&nbsp;&nbsp; =
\fi</div><div>\fi</div><div>%%</div><div>%\IfFileExists{txfonts.sty}{\Requ=
irePackage{txfonts}}{}</div><div>\IfFileExists{natbib.sty}{\RequirePackage=
{natbib}}{}</div><div>\bibpunct{(}{)}{,}{a}{}{,}</div><div>%\setlength{\bi=
bsep}{0pt}</div><div>\IfFileExists{fancyhdr.sty}{\RequirePackage{fancyhdr}=
}{}</div><div>\IfFileExists{tocbibind.sty}{\RequirePackage[nottoc,notlof,n=
otlot]{tocbibind}}{}</div><div>\IfFileExists{fontenc.sty}{\RequirePackage[=
T1]{fontenc}}{}</div><div>\IfFileExists{url.sty}{\RequirePackage{url}}{}</=
div><div>\IfFileExists{amssymb.sty}{\RequirePackage{amssymb}}{}</div><div>=
%\IfFileExists{amsmath.sty}{\RequirePackage{amsmath}}{}</div><div>\IfFileE=
xists{ifthen.sty}{\RequirePackage{ifthen}}{}</div><div>\IfFileExists{array=
sty}{\RequirePackage{array}}{}</div><div>\IfFileExists{multicol.sty}{\Req=
uirePackage{multicol}}{}</div><div>%%</div><div>\if@german</div><div>&nbsp=
;&nbsp; =
\IfFileExists{german.sty}{\RequirePackage{german}}{}</div><div>&nbsp;&nbsp=
; \def\captionsgerman{%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\acknowname{Danksagung}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\refname{Literaturverzeichnis}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\abstractname{Zusammenfassung}%</div><div>% &nbsp; &nbsp; =
\def\appendixname{Anhang}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\figurename{Abbildung}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\tablename{Tabelle}}</div><div>&nbsp;&nbsp; =
\selectlanguage{german}</div><div>\else</div><div>&nbsp;&nbsp; =
\def\captionsenglish{%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\acknowname{Acknowledgement}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\refname{References}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\abstractname{Abstract}%</div><div>% &nbsp; &nbsp; =
\def\appendixname{Appendix}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\figurename{Figure}%</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\def\tablename{Table}}</div><div>\fi</div><div>%%</div><div>\if@hype=
r\IfFileExists{hyperref.sty}{\RequirePackage{hyperref}}{}\fi</div><div>%%<=
/div><div>\DeclareRobustCommand*{\vect}[1]{\ensuremath{%</div><div>\mathch=
oice{\mbox{\boldmath$\displaystyle#1$}}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp; &nbsp; &nbsp; =
{\mbox{\boldmath$\textstyle#1$}}</div><div>&nbsp;&nbsp; &nbsp; &nbsp; =
&nbsp; &nbsp; {\mbox{\boldmath$\scriptstyle#1$}}</div><div>&nbsp;&nbsp; =
&nbsp; &nbsp; &nbsp; &nbsp; =
{\mbox{\boldmath$\scriptscriptstyle#1$}}}}</div><div>%%</div><div>\Declare=
RobustCommand*{\chem}[1]{\ensuremath{\mathrm{#1}}}</div><div>\DeclareRobus=
tCommand*{\unit}[1]{\def~{\,}\ensuremath{\mathrm{#1}}}</div><div>%%</div><=
div>\setcounter{topnumber}{2}</div><div>\def\topfraction{1.}</div><div>\se=
tcounter{bottomnumber}{1}</div><div>\def\bottomfraction{1.}</div><div>\set=
counter{totalnumber}{3}</div><div>\def\textfraction{0.}</div><div>\def\flo=
atpagefraction{.7}</div><div>\setcounter{secnumdepth}{3}</div><div>\setcou=
nter{tocdepth}{3}</div><div>%%</div><div>%\sloppy\clubpenalty4000\widowpen=
alty4000</div><div>%%</div><div>\def\town#1{\def\@town{#1}} &nbsp; =
&nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; =
\def\@town{???}</div><div>\def\refereea#1{\def\@refa{#1}} &nbsp; &nbsp; =
&nbsp; &nbsp; &nbsp; =
\def\@refa{???}</div><div>\def\refereeb#1{\def\@refb{#1}} &nbsp; &nbsp; =
&nbsp; &nbsp; &nbsp; =
\def\@refb{???}</div><div>\def\submitteddate#1{\def\@subdate{#1}} &nbsp; =
\def\@subdate{???}</div><div>\def\submittedyear#1{\def\@subyear{#1}} =
&nbsp; =
\def\@subyear{\number\year}</div><div>\def\examinationdate#1{\def\@exdate{=
#1}} =
&nbsp;\def\@exdate{???}</div><div>\def\publicationyear#1{\def\@pubyear{#1}=
} \def\@pubyear{\number\year}</div><div>\def\isbn#1{\def\@isbn{#1}} =
&nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; =
\def\@isbn{???}</div><div>%%</div><div>\def\maketitle{%</div><div>\begin{t=
itlepage}</div><div>\if@goe</div><div>&nbsp;&nbsp; =
\vspace*{20mm}</div><div>&nbsp;&nbsp; =
\begin{center}</div><div>&nbsp;&nbsp; =
{\Huge\bfseries\@title\par}</div><div>&nbsp;&nbsp; =
\vspace{30mm}</div><div>&nbsp;&nbsp; {\large =
Dissertation\\[1mm]</div><div>&nbsp;&nbsp; zur Erlangung des =
Doktorgrades\\[1mm]</div><div>&nbsp;&nbsp; der =
Mathematisch-Naturwissenschaftlichen =
Fakult\"aten\\[1mm]</div><div>&nbsp;&nbsp; der =
Georg-August-Universit\"at zu G\"ottingen</div><div>&nbsp;&nbsp; =
\vfill</div><div>&nbsp;&nbsp; vorgelegt =
von\\[1mm]</div><div>&nbsp;&nbsp; =
{\Large\bfseries\@author}\\[1mm]</div><div>&nbsp;&nbsp; aus =
\@town</div><div>&nbsp;&nbsp; \par\vspace{30mm}</div><div>&nbsp;&nbsp; =
G\"ottingen \space\@subyear}</div><div>&nbsp;&nbsp; =
\end{center}</div><div>\fi</div><div>\if@bs</div><div>&nbsp;&nbsp; =
\if@sub</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vspace*{10mm}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\begin{center}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;{\Huge\bfseries\@title\par}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vspace{20mm}</div><div>&nbsp;&nbsp; &nbsp; &nbsp;{\large Der =
Fakult\"at f\"ur Elektrotechnik, Informationstechnik, =
Physik\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;der Technischen =
Universit\"at Carolo-Wilhelmina\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;zu Braunschweig\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;zur =
Erlangung des Grades \if@fem einer \else eines \fi =
\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\if@fem Doktorin \else =
Doktors \fi der Naturwissenschaften\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;(Dr.rer.nat.)\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;eingereichte\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;Dissertation</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\par\vspace{30mm}</div><div>&nbsp;&nbsp; &nbsp; &nbsp;von =
\@author\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;aus =
\@town\par}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\end{center}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vfill</div><div>&nbsp;&nbsp; &nbsp; &nbsp;{\large 1. Referentin =
oder Referent: \@refa\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;2. =
Referentin oder Referent: \@refb\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;eingereicht am: \@subdate\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;m\"undliche Pr\"ufung (Disputation) =
am:\par}</div><div>&nbsp;&nbsp; \else</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vspace*{20mm}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\begin{center}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;{\Huge\bfseries\@title\par}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vspace{30mm}</div><div>&nbsp;&nbsp; &nbsp; &nbsp;{\large Von der =
Fakult\"at f\"ur Elektrotechnik, Informationstechnik, =
Physik\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;der Technischen =
Universit\"at Carolo-Wilhelmina\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;zu Braunschweig\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;zur =
Erlangung des Grades \if@fem einer \else eines \fi =
\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\if@fem Doktorin \else =
Doktors \fi der Naturwissenschaften\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;(Dr.rer.nat.)\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;genehmigte\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;Dissertation</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\par\vspace{30mm}</div><div>&nbsp;&nbsp; &nbsp; &nbsp;von =
\@author\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;aus =
\@town\par}</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\end{center}</div><div>&nbsp;&nbsp; =
\fi</div><div>\fi</div><div>\if@goeorbs\else\IfFileExists{firstpage.tex}{\=
include{firstpage}}</div><div>&nbsp;&nbsp; {Please provide file =
firstpage.tex}\fi</div><div>\end{titlepage}</div><div>\newpage % second =
page</div><div>\thispagestyle{empty}</div><div>\if@goe</div><div>&nbsp;&nb=
sp; \if@sub</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vspace*{150mm}</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\noindent =
{\large D7\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Referent: =
\@refa\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Korreferent: =
\@refb\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Tag der m\"undlichen =
Pr\"ufung: \par}</div><div>&nbsp;&nbsp; \fi</div><div>&nbsp;&nbsp; =
\if@print</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\noindent {\large =
{\bfseries Bibliografische Information Der =
Deutschen&nbsp;</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;Bibliothek}\\[3mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Die =
Deutsche Bibliothek verzeichnet diese Publikation in der =
Deutschen&nbsp;</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;Nationalbibliografie; detaillierte bibliografische Daten sind =
im&nbsp;</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Internet \"uber {\tt <a =
href=3D"http://dnb.ddb.de">http://dnb.ddb.de</a>} =
abrufbar.</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vfill</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\noindent =
D7\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Referent: =
\@refa\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Korreferent: =
\@refb\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Tag der m\"undlichen =
Pr\"ufung: \@exdate\\[30mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;ISBN =
\@isbn\\[5mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Copernicus =
Publications \@pubyear\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;<a =
href=3D"http://publications.copernicus.org">http://publications.copernicus=
org</a>\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\copyright\ =
\@author\\[5mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Printed in =
Germany\par}</div><div>&nbsp;&nbsp; \fi</div><div>&nbsp;&nbsp; =
\if@file</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\IfFileExists{secondpage.tex}{\include{secondpage}}</div><div>&nbsp;=
&nbsp; &nbsp; &nbsp;{Please provide file =
secondpage.tex}</div><div>&nbsp;&nbsp; =
\fi</div><div>\fi</div><div>\if@bs</div><div>&nbsp;&nbsp; =
\if@sub</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\null</div><div>&nbsp;&nbsp; =
\fi</div><div>&nbsp;&nbsp; \if@acc</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vspace*{150mm}</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\noindent =
{\large 1. Referentin oder Referent: =
\@refa\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;2. Referentin oder =
Referent: \@refb\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;eingereicht =
am: \@subdate\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;m\"undliche =
Pr\"ufung (Disputation) am: \@exdate\par}</div><div>&nbsp;&nbsp; =
\fi</div><div>&nbsp;&nbsp; \if@print</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\noindent {\large {\bfseries Bibliografische Information Der =
Deutschen&nbsp;</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;Bibliothek}\\[3mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Die =
Deutsche Bibliothek verzeichnet diese Publikation in der =
Deutschen&nbsp;</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;Nationalbibliografie; detaillierte bibliografische Daten sind =
im&nbsp;</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Internet \"uber {\tt <a =
href=3D"http://dnb.ddb.de">http://dnb.ddb.de</a>} =
abrufbar.</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\vfill</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\noindent 1. Referentin =
oder Referent: \@refa\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;2. =
Referentin oder Referent: \@refb\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;eingereicht am: \@subdate\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;m\"undliche Pr\"ufung (Disputation) am: =
\@exdate\\[30mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;ISBN =
\@isbn\\[5mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Copernicus =
Publications \@pubyear\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;<a =
href=3D"http://publications.copernicus.org">http://publications.copernicus=
org</a>\\[1mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;\copyright\ =
\@author\\[5mm]</div><div>&nbsp;&nbsp; &nbsp; &nbsp;Printed in =
Germany\par}&nbsp;</div><div>&nbsp;&nbsp; \fi</div><div>&nbsp;&nbsp; =
\if@file</div><div>&nbsp;&nbsp; &nbsp; =
&nbsp;\IfFileExists{secondpage.tex}{\include{secondpage}}</div><div>&nbsp;=
&nbsp; &nbsp; &nbsp;{Please provide file =
secondpage.tex}</div><div>&nbsp;&nbsp; \fi =
&nbsp;</div><div>\fi</div><div>\if@goeorbs\else\IfFileExists{secondpage.te=
x}{\include{secondpage}}</div><div>{Please provide file =
secondpage.tex}\fi</div><div>} % end of maketitle =
definition</div><div>%%</div><div>\newenvironment{cfig}{\begin{figure}[t]\=
begin{center}}{\end{center}\end{figure}}</div><div>%%</div><div>\def\chapt=
ername{}</div><div>\def\appendixname{}</div><div>%%</div><div>\pagestyle{f=
ancy}</div><div>\renewcommand{\chaptermark}[1]{\markboth{\thechapter\hskip=
2ex #1}</div><div>&nbsp;&nbsp; {\thechapter\hskip2ex =
#1}}</div><div>\renewcommand{\sectionmark}[1]{\markright{\thesection\hskip=
2ex =
#1}}</div><div>\fancyhf{}</div><div>\fancyhead[LE]{\nouppercase{\leftmark}=
}</div><div>\fancyhead[RO]{\nouppercase{\rightmark}}</div><div>\fancyfoot[=
LE,RO]{\thepage}</div><div>\renewcommand{\footrulewidth}{0pt}</div><div>\f=
ancypagestyle{plain}{\fancyhf{}\fancyfoot[LE,RO]{\thepage}%</div><div>&nbs=
p;&nbsp; =
\renewcommand{\headrulewidth}{0pt}\renewcommand{\footrulewidth}{0pt}}</div=
><div>%%</div><div>% avoids headings on otherwise empty pages before a =
new =
chapter</div><div>\def\cleardoublepage{\clearpage\if@twoside\ifodd\c@page\=
else</div><div>&nbsp;&nbsp; =
\hbox{}\thispagestyle{empty}\newpage</div><div>&nbsp;&nbsp; =
\if@twocolumn\hbox{}\newpage\fi\fi\fi}</div><div>%%</div><div>%\def\@schap=
ter#1{\addcontentsline{toc}{chapter}{#1} % for toc</div><div>% &nbsp; =
\@mkboth{#1}{#1} &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; =
&nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; &nbsp; % =
headings</div><div>% &nbsp; =
\if@twocolumn\@topnewpage[\@makeschapterhead{#1}]</div><div>% &nbsp; =
\else\@makeschapterhead{#1}\@afterheading\fi}</div><div>%</div><div>% =
problem: adds also toc to =
toc</div><div>%%</div><div>\def\@makechapterhead#1{\vspace*{50\p@}{\parind=
ent\z@\raggedright</div><div>&nbsp;&nbsp; =
\normalfont\interlinepenalty\@M\Huge\bfseries{%</div><div>&nbsp;&nbsp; =
\@hangfrom{\hskip\z@\relax\thechapter\hskip2ex}#1\@@par}</div><div>&nbsp;&=
nbsp; \par\nobreak\vskip40\p@}}</div><div>%%</div><div>%% End of file =
`thesis.cls'.</div><br></div></body></html>=

--Boundary_(ID_8vaiOHVH+3CE5q6BOg9/XQ)--
\fi

\errorcontextlines=9
\PassOptionsToPackage{demo}{graphicx}
\documentclass[braunschweig]{thesis}

\usepackage{caption}

%\usepackage[tight]{subfigure}
%\usepackage{epstopdf}
%\usepackage{fixltx2e}
%\usepackage{float}
%\usepackage{amsmath}
%\usepackage{txfonts}

\begin{document}

\section{Tracking the photospheric flows}

blablabla

\begin{figure*}[h]
\centering
\includegraphics[width=1 \textwidth]{./figures/photosphere1.jpeg}
\caption{Image of the solar photosphere in continuum with MDI/SoHO,  
with a close-up taken at higher resolution by FG-SOT/Hinode. The  
latter shows the granulation.}  \label{granulation}
\end{figure*}

\end{document}